DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2018-12-05 rebuttal.tex

\documentclass[12pt,a4paper]{article}
\usepackage{longtable}
\usepackage[LGR,T1]{fontenc}
\usepackage{lscape}
%\usepackage{pdfsync}
\usepackage{multirow}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{lastpage}
\usepackage{afterpage}
\usepackage{lettrine}
\usepackage{color,soul}
\usepackage[dvipsnames]{xcolor}
\usepackage{colortbl}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{titlesec}

%Palatino font
%\usepackage{pxfonts}
%\usepackage{libertine}
\usepackage[scaled=0.88]{beraserif}
\usepackage[scaled=0.85]{berasans}
\usepackage[scaled=0.84]{beramono}
\usepackage{mathpazo}
%\linespread{1.05}
\usepackage[T1,small,euler-digits]{eulervm}

\usepackage[nomessages]{fp}

 \definecolor{bleuUCLclair}{rgb}{.09, 0.569, 1}
\definecolor{bleuUCLfonce}{rgb}{ .13, .52, .86}
\definecolor{redBurn}{rgb}{.91, 0.29, 0.08}

\usepackage[colorlinks=true,urlcolor=redBurn,linkcolor=black]{hyperref}
\addtolength{\topmargin}{-1.5cm}
\addtolength{\textheight}{1.5cm}
\addtolength{\textwidth}{2cm}
\addtolength{\footskip}{2cm}
\setlength{\evensidemargin}{-0.5cm}
\setlength{\oddsidemargin}{-0.5cm}
\setlength{\arrayrulewidth}{0.25pt}

\renewcommand{\baselinestretch}{1.1} % Interligne

\newenvironment{maliste}%
{ \begin{list}%
	{\textcolor{bleuUCLfonce}{$\bullet$}\hspace{0.5cm}}%
	{\setlength{\labelwidth}{50pt}%
	 \setlength{\leftmargin}{25pt}%
	 \setlength{\itemsep}{30pt}}}%
{ \end{list} }

%\renewcommand{\headrulewidth}{0.0pt}
%\newcommand{\clearemptydoublepage}{%
%	\newpage{\pagestyle{empty}\cleardoublepage}}

%section like title in longtable
\newcommand{\seclong}[1]{\multicolumn{2}{@{}l}{{\Large\sffamily #1}} 
\vspace{0.5cm}
\\}

%enumerate on two columns
\newcounter{listlong}
\newcommand{\newlistlong}{\setcounter{listlong}{1}}
\newcommand{\iteml}[1]{%
\hspace{4.5cm}\textcolor{redBurn}{\arabic{listlong}}\stepcounter{listlong}%
&%
#1%

\\%
}

%left in column
\newcommand{\lcol}[1]{%
\begin{minipage}[t]{.35\textwidth}%

#1%

\end{minipage}%
}

\title{\vspace{-1cm}
\begin{flushleft} {\sffamily Rebuttal for paper \emph{Paper Number} }\end{flushleft}}
\date{\vspace{-1.7cm}\begin{flushleft}\sffamily DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications, Nathanaël Perraudin, Michaël Defferrard, Tomasz Kacprzak, Raphael Sgier \end{flushleft}}

\pagestyle{fancy} 
\fancyhf{}
\fancyfoot[R]{\sffamily\thepage\ / \pageref{LastPage}}
  \fancyfoot[L]{ }

\fancypagestyle{plain}{%
  \fancyhf{}%
  \fancyfoot[R]{\sffamily\thepage\ / \pageref{LastPage}}
  \fancyfoot[L]{ }
}

\renewcommand{\headrulewidth}{0.0pt}

\newcommand{\hlc}[2][yellow]{ {\sethlcolor{#1} \hl{#2}} }

\titleformat{\section}
  {\bfseries\scshape}{Reviewer \# \thesection}{1em}{}

\titleformat{\subsection}
  {\normalfont\scshape}{Comment \# \thesubsection}{1em}{}

\usepackage[framemethod=default]{mdframed}
%\mdfsetup{skipabove=\topskip,skipbelow=\topskip}

\global\mdfdefinestyle{comment}{%
     linecolor=red,linewidth=0.1cm,%
     leftmargin=-0.5cm,rightmargin=-0.5cm, innerleftmargin=0.4cm,innerrightmargin=0.4cm,
     topline=false,bottomline=false
}

\global\mdfdefinestyle{manuscript}{%
     linecolor=gray!20,linewidth=0.05cm,backgroundcolor=gray!20,%
     leftmargin=-0.5cm,rightmargin=-0.5cm, innerleftmargin=0.4cm,innerrightmargin=0.4cm
}

  \renewcommand{\subsectionautorefname}{Comment}

\newcommand{\nati}[1]{{\color[rgb]{.1,.6,.1}{NP: #1}}}
\newcommand{\TK}[1]{{\color{red}{TK: #1}}}
\newcommand{\todo}[1]{{\color[rgb]{.6,.1,.6}{TODO: #1}}}

\begin{document}
\maketitle

% Note: it is very convenient to refer to other comments using the referencing system of \LaTeX, such as here see  \autoref{comment:errorEq2}.

The authors would like to thank the reviewers for their comments, which significantly helped improving the manuscript. We believe we have addressed all the issues raised in the following rebuttal. 
%  For each comment, we have first highlighted the issue, then we provided an answer, and finally we described how the manuscript was adjusted.

\section*{Answer to the editor}

\begin{mdframed}[style=comment] 
Thank you for submitting your manuscript to Astronomy and Computing. I have received comments from reviewers on your manuscript. Your paper should become acceptable for publication pending suitable minor revision and modification of the article in light of the appended reviewer comments.  In particular, there is a request from a reviewer to have access to the data used in your analysis.

Though this is designated a minor revision, there are quite a few points that need to be addressed before the paper is acceptable for publication.  When resubmitting your manuscript, please carefully consider all issues mentioned in the reviewers' comments, outline every change made point by point, and provide suitable rebuttals for any comments not addressed.
\end{mdframed}

\todo{Answer to editor}

\section{}

\subsection{}

\begin{mdframed}[style=comment] 
I have read with great interest the paper by Perraudin and collaborators presenting a methodology (and the associated code) for performing deep machine learning on the sphere using novel algorithms.

The paper is well-written, and the results are clearly presented. The litterature as well as the advantages and limitations of new and existing algorithms are discussed thoroughly and fairly. I recommend this paper for publication, provided the following comments are addressed.
\end{mdframed}

\todo{Thanks the reviewer}

\subsection{}
\begin{mdframed}[style=comment] 
One fairly significant concern I have is how convincing the application to other machine learning algorithms is. The authors compare deep sphere to a SVM classifier, in the case of distinguising from two cosmological models from density maps. This classification is obviously a simplification of what would be done in real comsological analyses, but is sufficient for the purpose of demonstration here. However, the metrics adopted do not clearly highlight the superiority of DeepSphere. The SVM takes pixel histograms or power spectrum densities (PSDs) of each input map. It is not surprising that DeepSphere outperforms them given that it takes the whole map as input. Histograms and PSDs are very restrictive compressions of the data - dropping a lot of spatial information. An SVM or even a standard fully connected neural network (NN) would perform a lot better. It is unclear how that would compare to DeepSphere. But I would certainly expect DeepSphere to be a lot easier to train and have many fewer parameters, which I think is the key advantage that needs to be highlighted here. This is analogous to comparing a standard convolutional NN to a standard fully-connected NN. With appropriate depth both will extract the same information and perform equally well. But of course the CNN has fewer parameters and is easier to train since its structure naturally extracts spatial information. One wouldn't compare the CNN to a SVM trained on pixel histograms or PSDs, since those would not have access to the same information. It would be more natural to compare the CNN and the NN trained on the same maps, and examine the number of parameters, the time and number of examples needed for training to achieve the same performance, etc. This would truly highlight how the structure of the NN is a key advantage.
\end{mdframed}

\todo{General answer:
\begin{itemize}
  \item `An SVM or even a standard fully connected neural network (NN) would perform a lot better.'
  a) I think that SVM fails (I think this this an overfitting problem. This is the case at least for the noiseless case, TO CHECK)
  b) NN is computationally not even possible, or it would have to be a very striped down version of the network... We can actually try...
  \item The reviewer is correct with respect of the number of parameters
\end{itemize}
We should probably add to comparing curves SVM and fully connected
}

\subsection{}
\begin{mdframed}[style=comment] 
Page 4: "There is no known point set that achieves the analogue of uniform sampling in Euclidean space and allows exact and invertible discrete spherical harmonic decompositions of arbitrary but band-limited functions." This is ambigious. There exists multiple schemes for performing exact SHT on band-limited signals. Those sampling are more efficient than healpix - often in terms of number of pixels, and sometimes in terms of performance of the spherical harmonic transforms - and more accurate since the transform is exact (at least numerically - so typical errors are machine precision). Healpix transforms are only exact to 1e-7 when using iterative schemes, and that depends on Nside and lmax. Please clarify this paragraph. Obviously healpix is very advantageous for the pooling and other operations that take advantage of its hierarchical nature. This is well discussed in the paper - in particular how the fast graph algorithms could accomodate other sampling schemes.
\end{mdframed}

\todo{We have updated the manuscript to

\begin{mdframed}[style=manuscript] 
blabla
\end{mdframed}
}

\subsection{}
\begin{mdframed}[style=comment] 
"Our graph is constructed as an approximation of sphere S2, a 2D manifold embedded in R3, Indeed, [33] showed that the graph Laplacian converges to the Laplace-Beltrami when the number of pixels goes to infinity. While our construction will not exactly respect the setting defined by [33], we observe empirically strong evidence of convergence." This paragraph is unclear. Could you please add a few words to clarify what the approach of 33 is? 
\end{mdframed}

\todo{We added

\begin{mdframed}[style=manuscript] 
blabla
\end{mdframed}}

\subsection{}

\begin{mdframed}[style=comment] 
Figure 3: What was resolution used for those maps? Does the projection of the eigenvectors depend on the resolution and band-limit? 
\end{mdframed}

\todo{Answer}

\subsection{}
\begin{mdframed}[style=comment] 
Page 7: "That is much more efficient than filtering with spherical harmonics, even though HEALPix was de- signed as an iso-latitude sampling that has a fast spherical transform. That is especially true for smooth kernels which require a low polynomial degree K. Figure 4 compares the speed of low-pass filtering for Gaussian smoothing using the spherical harmonics and the graph-based method presented here. A naive implementation of our method is ten to twenty times faster for Nside = 2048, with K = 20 and K = 5, respectively, than using the spherical harmonic transform at lmax = 3Nside implemented by the highly optimized libsharp [38] library used by HEALPix." This is not so clear. Is the graph method also performing a Gaussian smoothing with the same characteristics? If so, what are those? How precise is the graph convolution approximating the true harmonic-space Gaussian convolution in this case? This is relevant since some approximations (for instance when analyzing cosmic microwave background data) require high-precision Gaussian smoothing. Finally, is lipsharp optimizing Gaussian smoothing in any way?
\end{mdframed}
\todo{Answer}

\section{}

\begin{mdframed}[style=comment] 
First of all, I would like to thank the authors for writing such a well thought out paper on such an interesting method.

The paper contains an extremely well described technique for performing convolutions on the sphere by first transforming the sphere into pixel space using graphs and then using graph-based convolutional filters that are radially symmetric. These convolutions are then used in neural networks, with an example showing classification of different cosmological scenarios directly from convergence maps. There are several benefits to this method over previously described neural networks on spheres (such as using spherical harmonics, which is slow, or using 2D CNNs, which has to learn to deal with distortions from the projection). I found the approximation of the graph approaching spherical harmonics extremely fascinating and will certainly looking more into this myself in the future!

I will happily recommend this paper for publication. Firstly though, I would like access to the data (privately rather than via emailing the group) so that I can check the notebooks thoroughly. The notebooks are extremely well written but I would like to run them first. As well as this I think there are a few extra applications which will improve the manuscript further, and make the method more appealing for users. I would recommend that these ideas are at least seriously considered by the authors and hopefully implemented: 
\end{mdframed}

\todo{Thank the reviewer + give him access to data}

\subsection{}

\begin{mdframed}[style=comment] 
The first, and most important is an example using masks. Since the authors mention how their method should in principle work well with masks, especially in comparison to HEALPix and Clebsh-Gordan transform spherical harmonic methods I think that it would be extremely enlightening to present this work, probably in the form of a performance vs. time plot. With this being done, the informational content of a masked version of the cosmological application could easily be included. This would, I'm sure, be attractive to the cosmological community as well as the ML community who are always looking for good ways to deal with masked data.
\end{mdframed}
\nati{I do not think we can practically compare with Clebsh-Gordan transform. We need to explain that to the reviewer. This is a sensitive point.}

\subsection{}

\begin{mdframed}[style=comment] 
Secondly, using only SVM as a comparison might skew how well this technique appears to works in a deep learning setting. In principle (and I say this not knowing for sure) 2D CNNs should probably be able to work as well as DeepSphere (although no where near as elegantly) by taking in a projected 2D image of the sky and using *loads* of filters to learn about the distortion at different parts of the projected sky. It would be interesting (and I think a fairly easy test) to check how well DeepSphere performs against a 2D CNN on projections. 
\end{mdframed}
\nati{This is something we should be abble to do. Actually, since we used 1/12 of the sphere, all the pixel grids we use are somehow squared. @Tomek, do you know if we can use directly that as a projection? Or do you know how to do the projection? What happen if the traditional 2D CNN is as good as our technique?}

\begin{mdframed}[style=comment] 
Similarly, I think it would useful to check how well the DeepSphere CNN does against the FCN when the input data is rotated randomly during training - I imagine that the results should match or probably exceed the FCN results and I think this would be a neat little test which doesn't involve needing extra data, just more augmentation.
\end{mdframed}
\nati{Problem here is that rotation is not simple on a sphere. So it will be very slow... In theory, I believe that CNN will match FCN. What shall we do?}

\subsection*{Minor comment}

\begin{mdframed}[style=comment] 
I should say, even though I made the above comments, I thought the paper was really excellent and well written. There were a couple of very minor points which I would like the authors to expand slightly on or consider.
\end{mdframed}

\subsection{}
\begin{mdframed}[style=comment] 
While our construction will not exactly respect the setting defined by [33], we observe empirically strong evidence of convergence.

- Could you add a plot comparing the graph Laplacian as a function of number of pixels and the Laplace-Beltrami operator so the reader can explicitly see the convergence?
\end{mdframed}

\todo{Very good question: but too complex answer for this paper. This is kept for a future contribution. + explain why}

\subsection{}

\begin{mdframed}[style=comment] 
We found out that the one proposed above works well for our purpose, and did not investigate other approaches, leaving it to future work.

- Are you really going to search through different weighting schemes in future work? I would be interested if you were, but could you speculate on what you expect to achieve by looking at different ones?
\end{mdframed}

\todo{Yes, it is likely that we are, + graph learning?}

\subsection{}
\begin{mdframed}[style=comment] 
Thanks to the softmax, the output $\bf{y}\in\mathbb{R}^{N_{classes}}$ of a neural network is a probability distribution over the classes, i.e., $y_i$ is the probability that the input sample belongs to class $i$.

- This is a common misconception which is incorrect and needs restating. It is *not* a probability distribution for the class, rather it is a discretised conditional distribution for the class given both the data and the weights and biases of the trained network. One way to see that the value of $y_i$ cannot be considered a probability (which implies P(y|x), i.e. the probability of a class given the data) is that if the network is untrained, would you retrieve back the correct probability? It's wholly dependent on the network.
\end{mdframed}
\nati{I totally agree with the reviewer, but I think this is only a formulation issue. By the way, this mean that the reviewer is knowledgable with ML, and may be Bayesian. He might be working on another spherical approach.}

\subsection{}
\begin{mdframed}[style=comment] 
For example, on images, the subject of interest is most often around the center of the picture. (Also... Contrast that with images, where the subject of interest is most often around the center of the picture.)

- I don't see that this is general at all, only if you are considering say, preprocessed postage stamps of galaxies. I can't think of many other cases in modern machine learning where the images contain the informative part only in the centre of an image. I do agree that the informative part is often localised though. Maybe this should be updated. 
\end{mdframed}
\todo{Answer}

\subsection{}
% Than can be seen as intrinsic data augmentation, as a CNN would need to see many rotated versions of the same data to learn the invariance.

%  - I think you mean "This" rather than "Than". I think, however, you are making a more important point here than you realise. It is not really "intrinsic data augmentation" but rather building a NN to suit your purposes. This should be done by the entire community as, at the moment, too many people blindly build their networks without considering their data and then struggle to get the best results out. The symmetries of the data should be the first thing to influence architecture of a network, exactly as you have reasoned. You might want to add a statement about how you know what your symmetries are and thus, you know how to build a well reasoned network which is more likely to work (and that other should do the same).

% As such, the SHT is only performed once on the input (no back and forth SHT between each layer). While this clever trick lowers the complexity of the convolution to $\mathcal{O}(N_{pix})$, the non-linearity is $\mathcal{O}(N^{3/2}_{pix})$.

%  - I don't understand what the difference is between the convolution and the "non-linearity". I think this sentence either needs explaining further. Is the process $\mathcal{O}(N_{pix})$ or $\mathcal{O}(N^{3/2}_{pix})$?

% When mentioning that you are 10 to 20 times faster than the SHT it should be noted that libsharp has amazing distributed mpi functionality which helps, although of course your method is genuinely lower order computationally (and can be distributed simply on a GPU).

% In Appendix A. (my personal favourite part) could you measure the percent difference that different eigenvectors differ from the spherical harmonics due to incomplete discretisation of the sphere, perhaps plotting side by side plots like Figure 3, but with SH too (or their differences). By the way, Figure A.12 is placed in the wrong section.

% In Appendix C. is there a connection between $t$ and $\sigma$ that you can make, or do you just fit them to be similar? If you just fit them to be similar I'm not sure where the relative differences come from. If they are related, however, then the relative differences make sense, but I can't see what the connection is.

% A question that I would like to know personally, are you thinking about a future upgrade to non-radial kernels?

% Finally, I noticed a few points in the text where there were typos or things which were difficult to follow and I have put a note of them below and tried to correct them as much as I can so that you don't have to put too much effort in - I can't wait to see this paper published. Sorry that the list is so extensive (I'm a bit of a stickler for details and trying to write papers concisely):

### 2018-12-07 scnn.tex
pre-screen reviewer 1

\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{eigenvectors}
	\caption{The first 16 eigenvectors of the graph Laplacian, an equivalent of Fourier modes, of a graph constructed from the HEALPix sampling of the sphere.
    Eigenvectors 1--3 could be associated with spherical harmonics of degree $\ell=1$ and order $|m|=(0,1)$, eigenvectors 4--8 with degree $\ell=2$ and order $|m|=(0,1,2)$, and eigenvectors 9--15 with degree $\ell=3$ and order $|m|=(0,1,2,3)$.
    Nevertheless, the graph eigenvectors are only approximating spherical harmonics.}
	\label{fig:graph_harmonics}
\end{figure}

@@

Let us consider the heat diffusion problem
\begin{equation}
  \L \b{f}(t) = \tau \partial_t \b{f}(t),
  \label{eqn:heat_equation}
\end{equation}
where $\b{f}: \R_+ \rightarrow \R^{N_{pix}}$. Given the initial condition
$\b{f}(0)$, the solution of~\eqnref{heat_equation} can be expressed as
\begin{equation*}
  \b{f}(t) = e^{-\L \tau t} \b{f}(0) = \U e^{-\bLambda t \tau} \U\trans \g{f}(0) = K_t(\L) \b{f}(0),
\end{equation*}
which is, by definition, the convolution of the signal $\b{f}(0)$ with the kernel $K_t(\lambda)=e^{-\tau t \lambda}$. Since the kernel $K_t$ is applied to the graph eigenvalues $\bLambda$, which can be interpreted as squared frequencies, it can also be considered as a generalization of the Gaussian kernel on the sphere. \figref{gaussian_filters_comparizon} shows the effect of the convolution by diffusing a unit of heat for $\tau=1$ at various times $t$. By comparing the graph convolution with the spherical symmetric Gaussian smoothing, we observe that both techniques lead to similar results (see \figref{gaussian_filters_comparizon}). While the graph convolution remains different from the spherical convolution, this small experiment shows that, providing the correct parameters, the graph convolution can approximate the spherical convolution. For additional insights about the graph kernel $K_t$, we refer to the \figref{gaussian_filters_visualization}.
% and~\ref{sec:filter_visualization}.

>>>>

\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{eigenvectors}
	\caption{The first 16 eigenvectors of the graph Laplacian, an equivalent of Fourier modes, of a graph constructed from the HEALPix sampling of the sphere \nati{($N_{side}=16$)}.
    Eigenvectors 1--3 could be associated with spherical harmonics of degree $\ell=1$ and order $|m|=(0,1)$, eigenvectors 4--8 with degree $\ell=2$ and order $|m|=(0,1,2)$, and eigenvectors 9--15 with degree $\ell=3$ and order $|m|=(0,1,2,3)$.
    Nevertheless, the graph eigenvectors are only approximating spherical harmonics.}
	\label{fig:graph_harmonics}
\end{figure}

@@

Let us consider the heat diffusion problem
\begin{equation}
  \tau \L \b{f}(t) = - \partial_t \b{f}(t),
  \label{eqn:heat_equation}
\end{equation}
where $\b{f}: \R_+ \rightarrow \R^{N_{pix}}$. Given the initial condition
$\b{f}(0)$, the solution of~\eqnref{heat_equation} can be expressed as
\begin{equation*}
  \b{f}(t) = e^{-\L \tau t} \b{f}(0) = \U e^{-\bLambda \tau t} \U\trans \g{f}(0) = K_t(\L) \b{f}(0),
\end{equation*}
which is, by definition, the convolution of the signal $\b{f}(0)$ with the kernel $K_t(\lambda)=e^{-\tau t \lambda}$. Since the kernel $K_t$ is applied to the graph eigenvalues $\bLambda$, which can be interpreted as squared frequencies, it can also be considered as a generalization of the Gaussian kernel on the sphere. \figref{gaussian_filters_comparizon} shows the effect of the convolution by diffusing a unit of heat for $\tau=1$ at various times $t$. By comparing the graph convolution with the spherical symmetric Gaussian smoothing, we observe that both techniques lead to similar results (see \figref{gaussian_filters_comparizon}). While the graph convolution remains different from the spherical convolution, this small experiment shows that, providing the correct parameters, the graph convolution can approximate the spherical convolution. For additional insights about the graph kernel $K_t$, we refer to the \figref{gaussian_filters_visualization}.
% and~\ref{sec:filter_visualization}.
